---
title: "ddp_pitch"
author: "Chuck Chekuri"
date: "4/4/2017"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(quantmod)
library(forecast)
library(knitr)
getSymbols(c("GOOG"))
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Plot the close prices

```{r getstock, echo = FALSE}
  	adjPrice <- GOOG[,6]
  	colnames(adjPrice) <- "Close"
  	plot(adjPrice, main=paste0("Close Prices for GOOG"))

```

## Show the prediction

```{r prediction, echo=FALSE}
        colnames(adjPrice) <- "Close"
  	fit  <- auto.arima(adjPrice)
  	preds <-forecast(fit, h=2)
  	kable(preds)
```

